---
title: "Confirmatory Factor Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
library(dplyr)
library(knitr)
library(corrplot)
library(psych)
library(ggfortify)
library(GPArotation)
library(pander)
library(ggplot2)
library(gplots)
library(gridExtra)
library(leaps)
library(pander)
library(stargazer)

parHiv <- read.delim("C:/Users/mitch/Documents/portfolio/data/parhiv.txt", header=TRUE, stringsAsFactors = FALSE, sep="\t")

x=parHiv$PB03
parHiv$PB03[x == 1] <- 4
parHiv$PB03[x == 2] <- 3
parHiv$PB03[x == 3] <- 2
parHiv$PB03[x == 4] <- 1

x=parHiv$PB07
parHiv$PB07[x == 1] <- 4
parHiv$PB07[x == 2] <- 3
parHiv$PB07[x == 3] <- 2
parHiv$PB07[x == 4] <- 1

x=parHiv$PB15
parHiv$PB15[x == 1] <- 4
parHiv$PB15[x == 2] <- 3
parHiv$PB15[x == 3] <- 2
parHiv$PB15[x == 4] <- 1

x=parHiv$PB21
parHiv$PB21[x == 1] <- 4
parHiv$PB21[x == 2] <- 3
parHiv$PB21[x == 3] <- 2
parHiv$PB21[x == 4] <- 1

x=parHiv$PB22
parHiv$PB22[x == 1] <- 4
parHiv$PB22[x == 2] <- 3
parHiv$PB22[x == 3] <- 2
parHiv$PB22[x == 4] <- 1

x=parHiv$PB25
parHiv$PB25[x == 1] <- 4
parHiv$PB25[x == 2] <- 3
parHiv$PB25[x == 3] <- 2
parHiv$PB25[x == 4] <- 1
```

I was given a data set where the researchers had an idea of how the variables would relate. These were nine factors about emotional health and our goal was to see if the variables would be grouped into these nine factors through confirmatory factor analysis. Only a couple were kind of close and the rest were far from it.
  
Since our goal was to see how the variables related, I first looked at how a computer would group the variables.

#Data Exploration for factor analysis

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
parHiv2 = parHiv %>% select(BSI01:BSI53) 
#parHiv2 = as.data.frame(
stanHiv = as.data.frame(scale(parHiv2))
#for(i in 1:53){
  #levels(parHiv2[,i]) = c("Not at all","A little bit","Moderately","Quite a bit","Extremely")
#}

corrplot(cor(na.omit(stanHiv)), tl.col="black",order = "hclust",addrect = 9)
extract_pca = princomp(na.omit(stanHiv))
var_pc <- (extract_pca$sdev)^2
qplot(x=1:length(var_pc), y=var_pc, geom=c("point", "line")) +
  xlab("PC number") + ylab("Eigenvalue") +  geom_hline(aes(yintercept=1))

```

The corrplot has some very small groupings of variables when it is forced to do 9. This may leave us with some lackluster factors that do not fit what is predicted. The scree plot tells us that we could take 2 factors if we followed the elbow rule or about 12 factors if we choose to go for an eigenvalue less than 1. We will still try 9 since that is what is expected. Now we will look at the factor loadings, given that there should be 9 factors. 

#Factors and their loadings

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
factors = principal(stanHiv, nfactors=9, rotate="varimax",scores=TRUE)

print(factors$loadings, cutoff = .40)

scores = data.frame(factors$scores)
```

Factor 1 (RC1): BSI03,04,05,09,12,15,22,50,53. BSI Obsessive compulsive subscale  
Factor 2 (RC4): BSI07,25,35,37,38,46,47,48,49,52,53. BSI Anxiety subscale  
Factor 3 (RC9): BSI20,21,24,41,42,43,51. BSI Interpersonal sensitivity subscale  
Factor 4 (RC5): BSI14,15,16,17,32. BSI Depression subscale  
Factor 5 (RC2): BSI04,06,10,13,40,41,46. BSI Hostility subscale  
Factor 6 (RC3): BSI01,02,07,23,29,30. Somatization subscale  
Factor 7 (RC7): BSI08,44,45. BSI Psychoticism subscale  
Factor 8 (RC6): BSI10,11,25,28 BSI Phobic anxiety subscale  
Factor 9 (RC8): BSI33,34. BSI Paranoid ideation subscale  

All of the factors match with at least one variable in the actual subscales except factor 9 which I am calling BSI Paranoid ideation even though it shares no factors. Factors 7 and 9 both share have a variable in Psychoticism but I chose Factor 7 simply because it came first. It turns out that things are not working as predicted. Now we will work on a model to predict the overprotection subscale.

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
names(scores) = c("BSI_obsessive","BSI_anxiety","BSI_interpersonal_sensitivity","BSI_depression","BSI_hostility","BSI_somatization","BSI_psychoticism","BSI_phobic_anxiety","BSI_paranoid_ideation")

parHivScores = data.frame(parHiv,scores)
```


```{r,include = FALSE, echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
paroverprotect = parHiv %>% select(PB03,PB07,PB08,PB09,PB10,PB13,PB15,PB19,PB20,PB21,PB22,PB23,PB25) %>% t() %>% data.frame()
overprotection = 1:252
for(i in overprotection){
  overprotection[i] = mean(paroverprotect[,i])
}
overprotection = data.frame(overprotection)

parHIV = data.frame(parHivScores,overprotection)
pander(summary(parHIV$overprotection))

```

#Data exploration for model building

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
p1 = ggplot(parHIV,aes(x=AGE,y=overprotection)) + geom_point() + geom_jitter()
parHIV$GENDER = factor(parHIV$GENDER)
p2 = ggplot(parHIV,aes(x=GENDER,y=overprotection)) + geom_boxplot()
parHIV$LIVWITH = factor(parHIV$LIVWITH)
levels(parHIV$LIVWITH) =c("both parents","one parent","other")
p3 = ggplot(parHIV,aes(x=LIVWITH,y=overprotection)) + geom_boxplot()
parHIV$SIBLINGS = factor(parHIV$SIBLINGS)
levels(parHIV$SIBLINGS) = c("no","yes")
p4 = ggplot(parHIV,aes(x=SIBLINGS,y=overprotection)) + geom_boxplot()
parHIV$HOWREL = factor(parHIV$HOWREL)
levels(parHIV$HOWREL) = c("very","somewhat","not at all")
p5 =  ggplot(parHIV,aes(x=HOWREL,y=overprotection)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4,p5)


```

From some bivariate graphs of demographic variables vs overprotection we can see that there may be a relationship between age and overprotection. Females may have a higher overprotection than males. Living situation may be something that effects overprotection. Both having siblings and religiousness do not seem very significant. 

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
parHIV$NGHB2 = factor(parHIV$NGHB2)
levels(parHIV$NGHB2) = c("not","somewhat","quite","very")
p6 = ggplot(parHIV,aes(x=NGHB2,y=overprotection)) + geom_boxplot() + ggtitle("drug addicts")
parHIV$NGHB5 = factor(parHIV$NGHB5)
levels(parHIV$NGHB5) = c("not","somewhat","quite","very")
p7 = ggplot(parHIV,aes(x=NGHB5,y=overprotection)) + geom_boxplot() + ggtitle("youth gang fights")
parHIV$NGHB6 = factor(parHIV$NGHB6)
levels(parHIV$NGHB6) = c("not","somewhat","quite","very")
p8 = ggplot(parHIV,aes(x=NGHB6,y=overprotection)) + geom_boxplot() + ggtitle("violence and violent crime")
parHIV$NGHB8 = factor(parHIV$NGHB8)
levels(parHIV$NGHB8) = c("not","somewhat","quite","very")
p9 = ggplot(parHIV,aes(x=NGHB8,y=overprotection)) + geom_boxplot() + ggtitle("drug dealing")
grid.arrange(p6,p7,p8,p9)
```

For the neighborhood variables it looks like drug addicts in the neighborhood and youth gang fights may effect overprotection. Violence and drug dealing to not seem significant. 

#Model building

##For this first model I will just use the BSI variables that maximize R^2 and some demographic variables. 

First we will see what automatic variable selection thinks. Then I will try out some models and end up with model 1.

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
regsubsets.out = regsubsets(overprotection~BSI_obsessive+BSI_anxiety+BSI_interpersonal_sensitivity+BSI_depression+BSI_hostility+BSI_somatization+BSI_psychoticism+BSI_phobic_anxiety+BSI_paranoid_ideation,data=parHIV)
summary.out <- summary(regsubsets.out)
plot(regsubsets.out, scale="adjr2")
max = which.max(summary.out$adjr2)
summary.out$which[3,]

model1 = lm(overprotection~AGE+LIVWITH+GENDER+BSI_interpersonal_sensitivity+BSI_hostility+BSI_paranoid_ideation, data = parHIV)
pander(summary(model1))
AIC(model1)

```



##For the second model I will try forward selection starting with age and then add neighborhood and bsi variables. 

Similarialy I made a second model.

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
model2 = lm(overprotection~AGE+BSI_hostility+BSI_somatization+NGHB2+NGHB8,data=parHIV)
pander(summary(model2))
AIC(model2)
```


  
Model 1 has a adjusted R^2 of .111 and a AIC of 361.1. Model 2 has an adjusted R^2 of .075 and an AIC of 373.02. Model 1 is a better model. For the final model we will add the effective variables in model 2 to model 1. 

```{r,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
fin_mod = lm(overprotection~AGE+LIVWITH+GENDER+BSI_hostility+BSI_paranoid_ideation+BSI_somatization+NGHB2+NGHB8, data = parHIV)

pander(summary(fin_mod))
AIC(fin_mod)

par(mfrow=c(2,2))
plot(fin_mod)
```

This model has an adjusted R^2 of .134 and an AIC of 360.509, making it the best model. The residuals of this model are mostly normal. For our coefficients we have -.04(age) which is a -.04 decrease in overprotection as age decreases. There is .144 increasing for living with one parent when compared to two parents and -.19 for other. There is -.1869 decrease in overprotection for males when compared to females. For BSI-hostility there is a .05 increase for each increase in the score. There is a -.01 decrease and a .032 increase for BSI_paranoid_ideation and BSI somatization respectively. For NGHB2, drug addicts in a neighborhood, we have .19 increase if it is somewhat a problem, .291 if it is quite a problem, and .397 if it is very much a problem. Finally for NGHB8, drug dealing, we have a .015 increase if it is somewhat a problem, -.214 decrease if it is quite a problem, and -.24 decrease if it is very much a problem.  













