---
title: "Fake News Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,  echo=FALSE, message = FALSE, warning = FALSE}
library(haven)
library(dplyr)
library(ggplot2)
library(magrittr)
library(gridExtra)
data = read_sav("C:/Users/mitch/Documents/portfolio/data/fakenews.sav")


pewdata <- data %$% data.frame(pew1 = factor(replace(pew1,pew1 == 9,NA),labels = c("Often","Sometimes","Hardly ever","Never")),
                               pew2 = factor(replace(pew2,pew2 == 9,NA),labels = c("Often","Sometimes","Hardly ever","Never")),
                               pew3 = factor(replace(pew3,pew3 == 9,NA),labels = c("Yes","No")),
                               pew4 = factor(replace(pew4,pew4 == 9,NA),labels = c("Yes","No")),
                               pew5a = factor(replace(pew5a,pew5a == 9,NA),labels = c("A great deal","A fair amount","Not much","None at all")),
                               pew5b = factor(replace(pew5b,pew5b == 9,NA),labels = c("A great deal","A fair amount","Not much","None at all")),
                               pew5c = factor(replace(pew5c,pew5c == 9,NA),labels = c("A great deal","A fair amount","Not much","None at all")),
                               pew6 = factor(replace(pew6,pew6 == 9,NA),labels = c("Very","Somewhat","Not very","Not at all")),
                               pew7 = factor(replace(pew7,pew7 == 9,NA),labels = c("Very","Somewhat","Not very","Not at all")),
                               sex = factor(sex,labels =c("Male","Female")),
                               age = as.numeric(age),
                               educ2 = factor(replace(educ2,educ2 == 9, NA),labels =c("Less than high school","High school incomplete","High school graduate","Some college","Associate degree","Bachelor degree","Some postgraduate","Postgraduate")),
                               relig = factor(replace(relig,relig == 99, NA),labels =c("Protestant","Roman Catholic","Mormon","Orthodox","Jewish","Muslim","Buddhist","Hindu","Atheist","Agnostic","Something else","Nothin in particulat","Christian","Unitarian")),
                               chr = factor(replace(chr, chr == 9, NA),labels =c("Yes","No")),
                               born = factor(replace(born, born == 9, NA),labels =c("Yes,world","No, would not")),
                               race = factor(replace(race,race == 9, NA),labels =c("White","Black or African American","Asian or Pacific Islander","Mixed race","Native American/American Indian","Other")),
                               inc = factor(replace(inc, inc == 99, NA),labels =c("Less than $10,000","$10,000 to under $20,000","$20,000 to under $30,000","$30,000 to under $40,000","$40,000 to under $50,000","$50,000 to under $75,000","$100,000 to under $150,000","$150,000 or over",NA)),
                               reg = factor(replace(reg, reg == 9, NA),labels=c(" Are you ABSOLUTELY CERTAIN that you are registered to vote at your current address","Are you PROBABLY registered, but there is a chance your registration has lapsed","Are you NOT registered to vote at your current address")),
                               party = factor(replace(party,party == 9, NA),labels=c("Republican","Democrat","Independent","No preference","Other party")),
                               partyln = factor(replace(partyln,partyln == 9, NA),labels = c("Republican","Democratic")),
                               ideo = factor(replace(ideo,ideo == 9, NA),labels=c("Very conservative","Conservative","Moderate","Libaral","Very libaral")))

```

```{r, include=FALSE, echo=FALSE, message = FALSE, warning = FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(ROCR)
library(caret)
library(questionr)
library(lmtest)
library(leaps)
library(pander)
library(ResourceSelection)
library(knitr)
```

Our goal for this project was to look at how different political parties viewed fakes news. Our data set was from the PEW Research center and it contained question such as "How well do you think you can determine a fake news story". These type of questions are the PEW questions which are:   
     
PEW1: How often do you come across news stories about politics that you think are not totally accurate?  
PEW2: And how often do you come across news stories about politics that you think completely made up?  
PEW3: Have you shared a news story you later found out was made up?  
PEW4: Have you shared a news story you thought at the time was made up?  
PEW5a: How much responsibility do members of the public have in trying to prevent made up stories from gaining public attention? 
PEW5b: How much responsibility do government and elected officials have in trying to prevent made up stories from gaining public attention? 
PEW5c: How much responsibility do social media websites have in trying to prevent made up stories from gaining public attention? 
PEW6: How confident are you in your ability to recognize news that is made up?  
PEW7: How much do you think these kind of stories leave Americans confused about the basic facts of current events?  
 

First we looked at how the political variables were distributed.  

## Political Affiliations of Population

```{r, fig.align='center', echo=FALSE, message = FALSE, warning = FALSE, out.width =1536, out.height=800}
party   <- ggplot(pewdata,aes(x=party, fill = ideo, group = ideo)) + 
  geom_bar(position = 'dodge') +
  geom_text(stat = 'count', position=position_dodge(width = 1),
            aes(y = ..count..,  label = ..count.., vjust = 0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = c(.7,.8)) + 
  guides(fill=guide_legend(ncol=2)) +
  scale_fill_manual(values=c('#BF6200','#FF8300',
                             '#9F008C','#416000',
                             '#AEFF00','#000000'),
                    name = 'Ideology',
                    na.translate=TRUE,
                    na.value='#000000') +
  ggtitle("Political Party")
party
```

Here we looked at political party and political ideology. A few of the responses were odd, with a couple very liberal republicans or very conservative democrats. 

## Pew questions vs Ideology

```{r, fig.align='center', echo = FALSE, message = FALSE, warning= FALSE,message=FALSE}

pew1ideo = pewdata %>% select(pew1,ideo) %>% na.omit()  
pew1ideotable = data.frame(prop.table(table(pew1ideo),margin=2))
p11= ggplot(pew1ideotable, aes(x=pew1,y=Freq,fill = ideo)) + geom_bar(position = "dodge",stat="identity") +
  geom_text(stat = 'count', position=position_dodge(width = 1),
            aes(y = ..count..,  label = ..count.., vjust = 0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = c(0.8, 0.7), legend.key.size = unit(.3,"line")) 

pew2ideo = pewdata %>% select(pew2,ideo) %>% na.omit()  
pew2ideotable = data.frame(prop.table(table(pew2ideo),margin=2))
p21 = ggplot(pew2ideotable, aes(x=pew2,y=Freq,fill = ideo)) + geom_bar(position = "dodge",stat="identity") +
  geom_text(stat = 'count', position=position_dodge(width = 1),
            aes(y = ..count..,  label = ..count.., vjust = 0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = c(0.8, 0.7), legend.key.size = unit(.3,"line"))

pew6ideo = pewdata %>% select(pew6,ideo) %>% na.omit()  
pew6ideotable = data.frame(prop.table(table(pew6ideo),margin=2))
p61 = ggplot(pew6ideotable, aes(x=pew6,y=Freq,fill = ideo)) + geom_bar(position = "dodge",stat="identity") +
  geom_text(stat = 'count', position=position_dodge(width = 1),
            aes(y = ..count..,  label = ..count.., vjust = 0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = c(0.8, 0.7), legend.key.size = unit(.3,"line"))
       

pew7ideo = pewdata %>% select(pew7,ideo) %>% na.omit()  
pew7ideotable = data.frame(prop.table(table(pew7ideo),margin=2))
p71 = ggplot(pew7ideotable, aes(x=pew7,y=Freq,fill = ideo)) + geom_bar(position = "dodge",stat="identity") +
  geom_text(stat = 'count', position=position_dodge(width = 1),
            aes(y = ..count..,  label = ..count.., vjust = 0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = c(0.8, 0.7), legend.key.size = unit(.3,"line"))

pewideo = grid.arrange(p11,p21,p61,p71) 
pewideo
```

Then we looked at how the PEW questions distributed across ideology. We can see that answers for the PEW questions do sometimes differ among ideologies. Specifically people who are very liberal or very conservative have stronger beliefs about PEW6. Now we will try to create a model.

```{r, echo=FALSE, message = FALSE, warning = FALSE}
levels(pewdata$ideo) = c("Extreme","Moderate" , "Moderate", "Moderate","Extreme")
pew2 <- pewdata %>% select(pew1,pew2,pew3,pew4,pew5a,pew5b,pew5c,pew6,pew7,ideo,sex,age,educ2) %>% na.omit

mod1 <- glm(ideo~.-ideo,data=pew2,family="binomial")
summary(mod1)
```

After selecting all of the PEW questions and a couple of extra variables we created a logistic model. The dependent variable here is political ideology. We changed the variable from 5 levels to just 2, extreme and moderate. Extreme included people who were very liberal and very conservative. Moderate was everything else. 

```{r, echo=FALSE, message = FALSE, warning = FALSE}
model.pred.prob <- predict(mod1, type='response')
pr <- prediction(model.pred.prob, mod1$y)
perf <- performance(pr, measure="tpr", x.measure="fpr")

perf.f1 <- performance(pr,measure="f")
plot(perf.f1)
```
With a glm we need to find a cutoff point for what group a probability goes into. This plot lets us get a good idea. The cutoff should be around when the line starts to drop off. I'm choosing .85. So, in this model a true positive for extreme will be less than .85 and a true negative for moderate will be greater than .85.

```{r, echo=FALSE, message = FALSE, warning = FALSE}

df = data.frame(model.pred.prob)
plot.mpp <- data.frame(prediction = model.pred.prob, 
                       truth = factor(mod1$y, labels=c("Extreme", "Moderate")))

ggplot(plot.mpp, aes(x=truth, y=prediction, fill=truth)) + 
      geom_jitter(width=.2) + geom_violin(alpha=.4) + theme_bw() + geom_hline(yintercept=.85)
```

This plot shows that we chose a good cutoff. For the moderates, most of them are above the line. If the line was any higher we would predict too many moderates falsely. 


```{r, echo=FALSE, message = FALSE, warning = FALSE}
plot.mpp$pred.class <- ifelse(plot.mpp$prediction <.85, 0,1)

plot.mpp$pred.class <- factor(plot.mpp$pred.class, labels=c("Extreme", "Moderate"))


confusionMatrix(plot.mpp$pred.class, plot.mpp$truth, positive="Extreme")
```

Our model had a 71% accuracy rate. The specificity (true negative rate) is ok at 74% and the sensitivity (true positive rate) is average at 53%. This could be because of the low sample size of extreme ideologies. Next we made a model with political party being the dependent variable. 

```{r, echo=FALSE, message = FALSE, warning = FALSE}
levels(pewdata$party)
levels(pewdata$party) = c("Republican","Democrat" , NA, NA,NA)
pew2 = pewdata %>% select(pew1,pew2,pew3,pew4,pew5a,pew5b,pew5c,pew6,pew7,party,sex,age,educ2) %>% na.omit
mod1 = glm(party~.-party,data=pew2,family="binomial")
mod1pp = predict(mod1,type = "response")
df = data.frame(mod1pp)

pr <- prediction(mod1pp, mod1$y)
perf <- performance(pr, measure="tpr", x.measure="fpr")
perf.f1 <- performance(pr,measure="f")
plot(perf.f1)
```

This time our cutoff will be .49. 

```{r, echo=FALSE, message = FALSE, warning = FALSE}
plot.mod1pp <- data.frame(prediction = mod1pp, 
                       truth = factor(mod1$y, labels=c("Republican", "Democrat")))

ggplot(plot.mod1pp, aes(x=truth, y=prediction, fill=truth)) + geom_jitter(width=.2) + geom_violin(alpha=.4) + theme_bw() + geom_hline(yintercept=.49)
```

This time republican is a positive and democrat is a negative. Again, we should have a higher specificity than sensitivity.  

```{r, echo=FALSE, message = FALSE, warning = FALSE}
plot.mod1pp$pred.class <- ifelse(plot.mod1pp$prediction <.49, 0,1)
plot.mod1pp$pred.class <- factor(plot.mod1pp$pred.class, labels=c("Republican", "Democrat"))

confusionMatrix(plot.mod1pp$pred.class, plot.mod1pp$truth, positive="Republican")

```

This model had a higher sensitivity than the previous one but it had a lower specificity. The overall accuracy was also lower.

Overall these models showed that accurate predictions should be possible with better data. I think with a higher sample size and more specific questions political ideology can be predicted with fake news opinions. 


