---
title: "KIR Analysis"
author: "Team Clueless"
output: html_document
---

```{r,echo=FALSE,message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
projectdata = read.delim("C:/Users/mitch/Documents/portfolio/data/youth29jul15.txt", header=TRUE,sep="\t")
```

```{r}
langa = as.factor(projectdata$LANGA)
levels(langa) = c("No", "Yes")
hl = as.factor(projectdata$HL)
levels(hl) = c("No", 'Yes')
friends=projectdata$friends_sexexp
con2 = as.factor(projectdata$CON2)
levels(con2) = c("Agree", 'Neither Agree or Disagree', "Disagree")
inter = as.factor(projectdata$intervention)
levels(inter) = c("No KIR", "KIR")
```

bivariate 1
```{r}
df4 = na.omit(data.frame(hl,friends))
ggplot(df4,aes(x = friends, col= hl)) + 
  geom_density() + ggtitle("Perceptions of Freinds Having Sex by Hispanic/Latino:") 
```
analysis: difference in mean

```{r}
t.test(friends~hl, data=df4)
```
There is strong evidence that Hispanic/Latino and non-Hispanic/Latino students have different perceptions about their friends having sex (p = 2.2e-16). 


bivariate 2
```{r}
df5= na.omit(data.frame(hl,langa))
ggplot(df5, aes(x = langa, fill = hl )) + geom_bar(position = "dodge")  + ggtitle("Speak English at Home")

addmargins(table(hl,langa))
round(prop.table(table(hl,langa),margin = 1),3)
```
analisys: difference in proportion?

```{r}
prop.test(x=c(714, 4462), n=c(778, 5525))
```
There is strong evidence that the proportion of students that do speak English at home is not the same for Hispanic/Latino and non-Hispanic/Latino students(p = 8.927e-14). 


bivariate 3
```{r}
ggplot(projectdata,aes(x = friends, col= inter)) + 
  geom_density() + ggtitle("Effects of KIR:") 
```
analysis: difference in mean

```{r}
df2 = na.omit(data.frame(inter,friends))
t.test(friends~inter, data=df2)
```
There is strong evidence that KIR did effect students perceptions of their friends having sex (p = 2.807e-12).



Multivariable model: We are looking to see if KIR had a different level of effect in students that are Hispanic or Latino. 
```{r}
model = aov(friends~inter + hl + langa)
summary(lm(model))


```

